---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# lcmr

<!-- badges: start -->
[![](https://img.shields.io/badge/devel%20version-0.3.0-blue.svg)](https://github.com/ykunisato/lcmr)
[![](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN checks](https://cranchecks.info/badges/summary/lcmr)](https://cran.r-project.org/web/checks/check_results_lcmr.html)
[![](https://travis-ci.org/ykunisato/lcmr.svg?branch=master)](https://travis-ci.org/ykunisato/lcmr)
<!-- badges: end -->

The goal of lcmr is to simulate phenomena of classical conditioning and fit the conditioning data with the latent cause model. 

lcmr is an R package of [Latent Cause Model: LCM]( https://github.com/sjgershm/LCM) and [Latent Cause Modulated Rescorla-Wagner model: LCM-RW](https://github.com/sjgershm/memory-modification) made by[Sam Gershman](http://gershmanlab.webfactional.com/people/sam.html).


## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` 
# install.packages("devtools")
devtools::install_github("ykunisato/lcmr")
```

## List of functions

- fit_lcm: Fit model to conditioning data
- infer_lcm: Simulate the classical conditioning with LCM
- infer_lcm_rw: Simulate the classical conditioning with LCM-RW


## How to simulate the renewal effect with LCM

The renewal effect is the phenomenon in which the change of context after extinction can lead to a return of conditioned response (CR). infer_lcm(X,opts) can simulate the renewal effect.

First, you have to prepare the design of conditioning. In followings, I prepare the experimental design consisted from acquisition (CS presents with US in context A(context = 1), 10 trials), extinction (CS presents without US in context B(context = 0),10 trials)  and test (CS without US in context A(context = 0), 10 trials).

```{r}
US <-      c(1,1,1,1,1,1,1,1,1,1,  0,0,0,0,0,0,0,0,0,0, 0,0,0,0,0,0,0,0,0,0)
CS <-      c(1,1,1,1,1,1,1,1,1,1,  1,1,1,1,1,1,1,1,1,1, 1,1,1,1,1,1,1,1,1,1)
Context <- c(1,1,1,1,1,1,1,1,1,1,  0,0,0,0,0,0,0,0,0,0, 1,1,1,1,1,1,1,1,1,1)
X <- cbind(US, CS, Context)
```

You can simulate the conditioned response using infer_lcm(). I simulate the conditioned response setting the alpha = 0.4 . The alpha is the concentrate parameter of the Chinese Restaurant Process.

```{r message=FALSE, warning=FALSE}
library(lcmr)
sim_res <- infer_lcm(X = X, opts = list(c_alpha = 1))
```

You can the plot of change of conditioned response through the trials using the following codes.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
sim_data <- data.frame(X, Trial = seq(1,length(US)), CR = sim_res$V)

sim_data %>% 
  ggplot(aes(x = Trial, y = CR)) +
  geom_line() +
  ylim(0,1)
```

You can draw the plot of the posterior probability of each cause using the following codes.

```{r}
sim_post <- as.data.frame(sim_res$post) 
sim_post %>% 
  mutate(Trial = seq(1,length(US))) %>% 
  rename(C01 = V1, C02 = V2, C03 = V3, C04 = V4, C05 = V5,
         C06 = V6, C07 = V7, C08 = V8, C09 = V9, C10 = V10) %>% 
  gather(key = "Cause", value = "post",-Trial) %>%
  mutate(Cause = as.factor(Cause)) %>% 
  ggplot(aes(x = Trial, y = post, color = Cause)) +
  geom_line() +
  ylim(0,1) +
  labs(y="Posterior probability")
```


## How to simulate the spontaneous recovery with LCM-RW

The spontaneous recovery is the phenomenon in which the time elapses following extinction can lead to a return of conditioned response (CR). infer_lcm_rw(X,opts) can simulate the spontaneous recovery.

First, you have to prepare the design of conditioning. In followings, I prepare the experimental design consisted from acquisition (CS1 presents with US, CS2 presents without US, 9 trials each CS), extinction (both CSs presents without US, 6 trials each CS)  and test (both CSs presents without US after 1 day, 3 trials each CS).

```{r}
US  <- c(1,0,0,1,1,0,1,0,1,1,0,0,1,0,0,1,1,0, 0,0,0,0,0,0,0,0,0,0,0,0, 0,0,0,0,0,0)
CS1 <- c(1,0,0,1,1,0,1,0,1,1,0,0,1,0,0,1,1,0, 1,1,0,1,0,0,1,0,1,1,0,0, 1,0,0,1,1,0)
CS2 <- c(0,1,1,0,0,1,0,1,0,0,1,1,0,1,1,0,0,1, 0,0,1,0,1,1,0,1,0,0,1,1, 0,1,1,0,0,1)
time <- c(0, seq(1:17)*4, 100+seq(1:12)*4, 86400+seq(1:6)*4)
X <- cbind(time, US, CS1, CS2)
```

You can simulate the conditioned response using infer_lcm(). I simulate the conditioned response setting the alpha = 0.45 and the eta = 0.2. The alpha is the concentrate parameter of the Chinese Restaurant Process and eta is the learning rate of the RW model.

```{r message=FALSE, warning=FALSE}
library(lcmr)
sim_res <- infer_lcm_rw(X = X, opts = list(c_alpha = 0.45, eta = 0.2))
```

You can the plot of change of conditioned response through the trials using the following codes.

```{r message=FALSE, warning=FALSE}
library(tidyverse)
sim_data <- data.frame(X,
                       Trials_cs1 = cumsum(CS1),
                       Trials_cs2 = cumsum(CS2),
                       CR = sim_res$V)

sim_data %>% 
  filter(CS1 == 1) %>% 
  ggplot(aes(x = Trials_cs1, y = CR)) +
  geom_line() +
  ylim(0,1) +
  labs(x = "Trial", y = "CR of CS1")

sim_data %>% 
  filter(CS2 == 1) %>% 
  ggplot(aes(x = Trials_cs2, y = CR)) +
  geom_line() +
  ylim(0,1) +
  labs(x = "Trial", y = "CR of CS2")
```


## How to fit LCM to the conditioning data.

You have to prepare the data as long format containing the following variables (Order and name is exactly the same as following):

- ID  Subject ID
- CR  Conditioned Response
- US  Unconditioned Stimulus
- CS  Conditioned Stimului or Context.  If using multiple CS, set variables name as CS1,CS2,CS3...

I make the synthetic data for model fitting with the experiment design of the renewal effect.

```{r}
US <-      c(1,1,1,1,1,1,1,1,1,1,  0,0,0,0,0,0,0,0,0,0, 0,0,0,0,0,0,0,0,0,0)
CS <-      c(1,1,1,1,1,1,1,1,1,1,  1,1,1,1,1,1,1,1,1,1, 1,1,1,1,1,1,1,1,1,1)
Context <- c(1,1,1,1,1,1,1,1,1,1,  0,0,0,0,0,0,0,0,0,0, 1,1,1,1,1,1,1,1,1,1)
X <- cbind(US, CS, Context)
```


I make synthetic data for 100 participants.

```{r}
number_of_perticiapnts <- 100
participants_alpha <- runif(number_of_perticiapnts, 0, 10)
data <- NULL
for (i in 1:number_of_perticiapnts) {
  sim_data <- infer_lcm(X, opts = list(c_alpha = participants_alpha[i]))
  sim_df <- data.frame(ID = rep(i,length(US)), CR = sim_data$V, X)
  data <- rbind(data,sim_df)
}
```

You can estimate parameter alpha using fit_lcm(data, model, opts, parameter_range, parallel, estimation_method). You have to specify the following argument for model fitting:

- model: 1 = latent cause model, 2 = latent cause modulated RW model

- parameter_range:  range of parameter(a_L, a_U, e_L, e_U)

- stimation_method:  0 = optim or optimize(lcm), 1 = post mean(only latent cause model)

```{r}
results <- fit_lcm(data, 
                   model = 1, 
                   parameter_range = list(a_L = 0, a_U = 15),
                   estimation_method = 0)
```

I check the parameter recovery. It looks well recovery of parameter alpha.

```{r}
parameter_recovery <- data.frame(true_alpha = participants_alpha, estimated_alpha = results$parameters$alpha)
parameter_recovery %>% 
  ggplot(aes(x = true_alpha, y= estimated_alpha)) +
  geom_point()
```

## How to fit LCM-RW to the conditioning data.

The fit_lcm() allow to estimate alpha and eta of LCM-RW (set model = 2). However, fit_lcm() can not to estimate parameter adequatly at this time(fit_lcm can not recover the parameter adequately).


## Bugs and question

Please report on this repository's [issues](https://github.com/ykunisato/lcmr/issues)
